### 2022-06-27 response-letter.md

Dear editor,

Thank you for the valuable feedback on our manuscript. We are pleased that you
and the reviewer found interest and worth in our manuscript. We addressed each comment and found that the manuscript is indeed clearer now, and we hope to have it approved after your consideration.

We have added more context in the introduction section, but we've found it hard
to expand the first paragraph. We believe that this particular paragraph needs
to be more succinct, at the same time that it provides a complete picture of the
paper, framed in a storytelling structure. Then, on the second paragraph, we
start providing literature background and many examples, to help the reader
understand where our paper stands in the scientific context. We have included
discussions about the quality and accuracy of range maps (lines XX-XX), examples
of data biases (lines XX-XX), and we addressed the uniqueness of our study
compared to previous research (lines XX-XX). We hope that this structure is
clarifying and appealing to the readers.

Regarding the GBIF data cleaning, retrieval and the validation steps, we added a
more detailed description of our process (lines XX-XX, XX-XX and XX-XX). Data
retrieving is detailed in the source code provided on our GitHub repository
(code/06-GBIF_data.jl). We used the list of species from the network dataset and
retrieved any occurrence within our bounding box (between longitudes -20.0 and
55.0 degrees; and latitudes -35.0 and 40.0 degrees). Some taxonomic
investigation needed to be done due to inconsistencies between GBIF and IUCN
nomenclatures (in the case of _Taurotragus oryx_). We didn't perform any further
data cleaning procedures because this could hinder the goals of the paper, which
was to investigate the characteristics of the data available for research. As we
needed to compare two sets of raw data, and data cleaning protocols are
dependent on the use objectives, we chose to avoid data manipulation both for
GBIF and IUCN datasets.

For the validation step, we understand our phrasing was confusing, and we have
rewritten this section to clarify our method. Rather than using
absence data, what we did was to transform the point data into raster files, and
then restrict our range maps to the locations where we had GBIF occurrences.
Therefore, the pixels where there are no GBIF occurrences are not considered
true absence of a species, but absence of **data** about the occurrence of that
species.

Finally, we have addressed all the other minor comments throughout the text as
described below:
- We've added citations on the Methods section to support our argument that a
  mismatch between ranges can be originated from different sources (L. 68-70)
- We have suppressed lines 214-217.
- The abstract is now restructured according to PeerJ guidelines.
- References to tables and figures in the Discussion section were suppressed,
  with the exception of Figure 05 (line XX), which is an example of our results
  to clarify an important point of the discussion.
- Lines 22-23 (now lines XX-XX): the cited literature are not examples of models
  that do not take ecological interactions into account, but studies that have
  demonstrated how ecological interactions might be responsible for shaping a
  species' distribution. We've rewritten the sentence to clarify that.
-  Lines 68-70: we have rewritten the sentence to include the consideration
   misestimation of both species in a trophic interaction. We have also added
   citations and examples.
- We have addressed the grammatical suggestions throughout the text, except for
  the legend of figure 5. We thought that the suggested phrasing adds a little
  bit of confusion to the comprehension of the figure.

We hope our revised manuscript is now more appropriate for publishing at PeerJ.
Reiterating our appreciation for yours and the reviewer's thoughtful comments,
we are available to answer any further questions.

Best regards,
The Authors

### 2022-07-08 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@7dfa55f42d368132ed2c1af3d1475ee71992276e 🚀

\vfill
\textbf{\sffamily Abstract: }Range maps are a useful tool to describe
the spatial distribution of species. However, they often need to be used
with caution, as they essentially represent a rough approximation of a
species' suitable habitats. When stacked together, the resulting
communities in each grid cell may not always be realistic, especially
when species interactions are taken into account. Here we show the
extent of the mismatch between range maps, provided by the International
Union for Conservation of Nature, and species interactions data. More
precisely, we show that local networks built from those stacked range
maps often yield unrealistic communities, where species of higher
trophic levels are completely disconnected from primary producers. We
use the well-described Serengeti food web of mammals and plants as our
case study, and provide updated range maps that take into account
food-web structure. In our analysis, most predator ranges were
restricted by the absence of herbivores. This restriction was sometimes
contradicted by GBIF occurrences, suggesting the mismatch can be due
either to the lack of information about ecological interactions or about
the geographical occurrence of preys. We finally discuss general
guidelines to help identify defective data among distributions and
interactions data, and we recommend this method as a valuable way to
assess weather the occurrence data that are being used, even if
incomplete, is ecologically accurate.
\vfill

@@

Finding a species in a certain location is like finding an encrypted
message that travelled through time. It carries the species'
evolutionary history, migration patterns, as well as any direct and
indirect effects generated by other species (some of which we may not
even know exist). Ecologists have been trying to decode this message
with progressively more powerful tools, from their own field notes to
highly complex computational algorithms. However, to succeed in this
challenge it is important to have the right clues in hand. There are
many ways we can be misled by data - or the lack of it: taxonomic
errors, geographic inaccuracy, or sampling biases (Ladle and Hortal
2013; Hortal et al. 2015; Poisot et al. 2021). One way to identify - and
potentially fix - these errors is to combine many different pieces of
information about the occurrence of a species, so agreements and
mismatches can emerge. Here we suggest jointly analysing species
occurrence (range maps and point occurrences) and ecological
interactions to identify mismatches between datasets.

@@

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (2021). These maps can be used in macroecological inferences in
the lack of more precise information (Fourcade 2016; Alhajeri and
Fourcade 2019), but it has been recommended that they are used with
caution since they tend to underestimate the distribution of species
that are not well-known (Herkt, Skidmore, and Fahr 2017), do not
represent spatial variation in species occurrence and abundance (Dallas,
Pironon, and Santini 2020), and can include inadequate areas within the
estimated range. Another source of species distribution information is
the Global Biodiversity Information Facility (GBIF), which is an online
repository of georeferenced observational records that come from various
sources, including community science programs, museum collections, and
long-term monitoring schemes. A great source of bias in these datasets
is the irregular sampling effort, with more occurrences originated from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods and
data aggregation (Poisot et al. 2020; Hortal et al. 2015). Nevertheless,
we have witnessed an increase in the availability of biodiversity data
in the last decades, including those collected through community science
projects (Callaghan et al. 2019; Pocock et al. 2015) and dedicated
databases, such as Mangal (Poisot et al. 2016). This provides an
opportunity to merge species distribution and ecological interaction
data to improve our predictions of where a species may be found across
large spatial scales (e.g., continental and global).

@@

We investigated the mismatch between savannah species ranges and
interactions in Africa (fig.~\ref{fig:richness}). These ecosystems host
a range of different species, including the well characterized
predator-prey dynamics between iconic predators (e.g., lions, hyenas and
leopards) and large herbivores (e.g., antelopes, wildebeests and
zebras), as well as a range of herbivorous and carnivorous small
mammals. The Serengeti ecosystem has been extensively studied and its
food web is one of the most complete we have to date, including primary
producers identified to the species level. Here we focus on six groups
of herbivores and carnivores from the Serengeti Food Web Data Set
(Baskerville et al. 2011). These species exhibit direct antagonistic
(predator-prey) interactions with one another and are commonly found
across savannah ecosystems on the African continent (McNaughton 1992).
Plants in the network were included indirectly in our analyses as we do
not expect the primary producers to significantly influence the range of
herbivores for several reasons. Firstly, many savannah plants are
functionally similar (i.e., grasses, trees and shrubs) and cooccur
across the same habitats (Baskerville et al. 2011). Secondly, herbivores
in the network are broadly generalists feeding on a wide range of
different plants across habitats. Indeed, out of 129 plants in our
dataset, herbivores (n = 23) had a mean out degree (mean number of
preys) of around 22 (std = 17.5). There is also an absence of global
range maps for many plant species (Daru 2020), which prevents their
direct inclusion in our analysis. Therefore, we assume that plants
consumed by herbivores are present across their ranges, and as such the
ranges of herbivores are not expected to be significantly constrained by
the availability of food plants.

From the wider ecological network presented in Baskerville (2011), we
sampled interaction data for herbivores and carnivores. This subnetwork
contained 32 taxa (23 herbivores and 9 carnivores) and 84 interactions,
and had a connectance of 0.08. Although self-loops are informative, we
removed these interactions to allow for the original IUCN ranges of
predators with cannibalistic interactions to be adjusted. We treated
this overall network as a metaweb since it \emph{should} contain all
potential species interactions between mammalian taxa occurring across
savannah ecosystems such as the Serengeti.

We compiled IUCN range maps for the 32 species included in the metaweb
from the Spatial Data Download portal
(www.iucnredlist.org/resources/spatial-data-download), which we
rasterized at 10 arc-minute resolution (\textasciitilde19 km² at the
equator). We then combined interaction data from the metaweb and
cooccurrence data generated from species ranges to create networks for
each raster pixel. This generated a total of 84,244 pixel-level
networks. These networks describe potential predation, not actual
interactions: the former is derived information from the metaweb, and
the later is contingent to the presence of herbivores.

@@

We calculated the geographical overlap, i.e.~the extent to which
interacting predator and prey species cooccurred across their ranges, as
\(a/(a + c)\), where \(a\) is the number of pixels where predator and
prey cooccur and \(c\) is the number of pixels where only the focal
species occur. This index of geographical overlap can be calculated with
prey or predators as the focal species. Values vary between 0 and 1,
with values closer to 1 indicating that there is a large overlap in the
ranges of the two species and values closer to 0 indicating low
cooccurrence across their ranges. For each predator species we
calculated its generality to understand whether the level of trophic
specialization (i.e., number of prey items per predator) affects the
extent to which the ranges of the species were altered. One would assume
that predators with a greater number of prey taxa (i.e., a higher
generality) are less likely to have significant changes in their range
as it is more likely that at least one prey species is present across
most of their range.

@@

For each species in the dataset we collated point observation data from
GBIF (www.gbif.org), and condensed these data into pixels representing
presence or absence of the focal taxon. These data were used to validate
the range adjustments made based on species interactions (see previous
section). To do so, we calculated the proportion of GBIF presence pixels
occurring within both the original and adjusted species ranges. We then
compared these proportions for the predators to verify if the range
adjustments removed locations with GBIF observations, hence likely true
habitats.

@@

Predators with less preys lose more range with our method
(fig.~\ref{fig:degree}). For instance, both \emph{Leptailurus serval}
and \emph{Canis mesomelas} have only one prey in the Serengeti food web
(tbl.~\ref{tbl:everyone}), each of them with a very small range compared
to those of their predator. This discrepancy between range sizes
promotes significant range loss. On the other hand, predators of the
genus \emph{Panthera} are some of the most connected species, and they
also lose the least proportion of their ranges. This mismatch between
predators and preys can also be a result of taxonomic disagreement
between the geographical and ecological data. Although \emph{Canis
aureus} has the same number of prey as \emph{Caracal caracal}, none of
the prey taxa of the former occurs inside its original range
(tbl.~\ref{tbl:everyone}), which results in complete range loss.

@@

There was high variation in the overlap of predator and prey ranges
(fig.~\ref{fig:geo_diss}). The high density of points on left-hand side
of fig.~\ref{fig:geo_diss} indicates that most preys have small ranges
in comparison to those of the set of carnivores in the networks,
resulting in either low overlap between both ranges (bottom) or high
overlap of ranges because much of that of the prey is within predators'
range (top). The top-right side of the plot encompasses situations where
the ranges of both predator and prey are similar and overlapping, while
the bottom-right part of the plot represents a situation where the range
of the predator is smaller than that of its prey and much of it occurs
within the preys' range. For example, \emph{Panthera pardus} had many
preys occurring inside its range, with highly variable levels of overlap
(tbl.~\ref{tbl:everyone}). In general, species exhibited more consistent
values of prey-predator overlap, than predator-prey overlap -- indicated
by the spread of points along the x-axis, yet more restricted variation
on the y-axis (fig.~\ref{fig:geo_diss}). There was also no overall
relationship between the two metrics, or for any predator species.

@@

degrees, total original range size (in pixels) and proportion of their

@@

(\emph{Loxodonta africana}) and how \emph{Canis aureus}'s range do not

@@

The proportion of GBIF pixels in updated ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Four species showed no difference of proportion while three species
showed only small differences (proportions of 0.01 to 0.05). On the
other hand, two species, \emph{Canis mesomelas} and \emph{Leptailurus
serval}, showed very high differences, with overlaps lower by 0.548 and
0.871 respectively. For \emph{Leptailurus serval}, none of the GBIF
observations occurred in the updated range. These two species are also
the only predators with a single prey in our metaweb. Our results
delineate how a mismatch between GBIF and IUCN databases differ greatly
with small changes in herbivore species ranges, and it is somewhat
positively related to range size for predator species. Moreover, we show
that accounting for interactions does not necessarily aggravates this
dissimilarity, but it is relevant for species with little ecological
information or specialists.

@@

The jackal is a widespread taxon in northern Africa, Europe and
Australasia, generally well adapted to local conditions due to its
largely varied diet (Tsunoda and Saito 2020; Krofel et al. 2021).
Because of that, we expected that the \emph{Canis} species in our
dataset would be the ones losing the least amount of range, with a
higher value of the proportion of GBIF pixels within their IUCN range
maps. However, the taxonomy of this group is a matter of intense
discussion, as molecular and morphological data seem to disagree in the
clustering of species and subspecies (Krofel et al. 2021; Stoyanov
2020). This debate is indeed reflected in our analysis: the GBIF
identification of the golden jackal is incompatible with the one used by
IUCN, each of them mapping its distribution in completely different
places. This led to a complete exclusion of \emph{Canis aureus} from its
original range in our analysis, despite the fact that this species has
four documented preys in our metaweb. This example illustrates how the
taxonomic, geographical and ecological data can be used to validate one
another.

Here we show that when ecological interaction data (predator-prey
interactions within food webs) are used to refine species range maps,
there are significant reductions in the IUCN range size of predatory
organisms. Despite showing the potential importance in accounting for
species interactions when estimating the range of a species, it remains
unclear the extent to which the patterns observed represent ecological
processes or a lack of data. In the following sections we discuss the
implications of our findings, in terms of species range maps,
interaction data and the next steps required to enhance understanding of
species distributions using information on ecological networks.

@@

In the Serengeti food web there is a positive relationship between the
out degrees of predators and the size of their ranges
(tbl.~\ref{tbl:everyone}). In addition, our results show that there is a
negative relationship between the relative loss of predators' ranges and
their number of preys (fig.~\ref{fig:degree}), reinforcing the idea that
generalist species can preserve their distributions longer while losing
interactions. The factors limiting the geographical range of a species
in a community can vary with connectivity and richness (Svenning et al.
2014). Younger communities may be more affected by environmental
limitations because they are dominated by generalist species, while
older metacommunities are probably affected in different ways in the
centre of the distribution, at the edge of ranges and in sink and source
communities (Svenning et al. 2014; Godsoe et al. 2017; Cazelles et al.
2016; Bullock et al. 2000). Additionally, it is likely that species with
larger ranges of distribution and those that are more generalists would
co-occur with a greater number of other species (Dáttilo et al. 2020),
while dispersal capacity of competitive species modulate their
aggregation in space and the effect of interactions on their range
limits (Godsoe et al. 2017).

@@

The geographical mismatch between predators and preys have ecological
consequences such as loss of ecosystem functioning and extinction of
populations (Anderson et al. 2016; Dáttilo and Rico-Gray 2018; Pringle
et al. 2016; Young et al. 2013). Climate change is one of the causes of
this, leading, for instance, to the decrease of plants populations due
to the lack of pollination (Bullock et al. 2000; Afkhami, McIntyre, and
Strauss 2014; Godsoe et al. 2017). However, this mismatch can also be
purely informational. When the distribution of predators and preys do
not superpose, it can mean we lack information about the distribution of
either species or about their interactions (e.g., predators may be
feeding on different species than the ones in our dataset outside the
Serengeti ecosystem). Here we addressed part of this problem by
comparing the IUCN range maps with GBIF occurrences, which helped us
clarify what is the shortfall for each species.

The lack of superposition between IUCN range maps and GBIF occurrences
suggests that we certainly do miss geographical information about the
distribution of a certain species, but it is not an indicative about the
completeness of the information about ecological interactions. However,
if both GBIF and IUCN occurrences tend to superpose and still the
species is locally removed, this indicates we don't have information
about all its interactions. The combination of this rationale with our
method of updating range maps based on ecological interactions allows us
to have a clearer idea of which information we are missing. For example,
the lion (\emph{Panthera leo}) was one of the species with the smallest
difference between the original and the updated ranges
(fig.~\ref{fig:degree}), but 59.5\% of the GBIF occurrences for this
species fell outside the IUCN range (fig.~\ref{fig:gbif}). In this
particular case, the IUCN maps seem to agree with species interactions
data. However, the disagreement between the IUCN and the GBIF databases
is concerning and suggests that the IUCN maps might underestimate the
lion's distribution. On the other hand, \emph{Leptailurus serval} and
\emph{Canis mesomelas} are two of the three species that lose the higher
proportion of range due to the lack of paths to a herbivore
(fig.~\ref{fig:degree}), but are also some of the species with the
higher proportion of GBIF occurrences inside IUCN range maps
(fig.~\ref{fig:gbif}). This indicates that the information we are
missing for these two species is related to either the occurrence of an
interaction or the presence of interacting species. To illustrate that,
we mapped the GBIF data for the prey of \emph{Leptailurus serval}, with
a mobility buffer around each point (fig.~\ref{fig:serval}). When
considering GBIF data, approximately 53\% of the prey's occurrences are
within the portion of the serval's range that was lost. With the buffer
area, this corresponds to 13\% of the lost range. This means that by
adding GBIF information, we would reduce the loss of range (or
information) for the predator by 13\% since its distribution is
conditional on the occurrence of its preys.

\begin{figure}
\hypertarget{fig:serval}{%
\centering
\includegraphics{figures/serval_mismatch_combined.png}
\caption{Mismatch between serval's range loss and GBIF occurrence of its
prey. The left panel shows the reduction of serval's range when we
consider the IUCN data on its prey. On the right panel, we added GBIF
data on both serval and its prey, with a buffer for the prey to account
on species mobility.}\label{fig:serval}
}
\end{figure}

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. GH,
FB, GD, and NF are funded by the NSERC BIOS\(^2\) CREATE program; FB,
NF, and TP are funded by IVADO; NF and TP are funded by a donation from
the Courtois Foundation; GD is funded by the FRQNT doctoral scholarship;
TP is funded by the Canadian Institute of Ecology \& Evolution; FW is
funded by the Royal Society (Grant number:
CHL\textbackslash R1\textbackslash180156).

>>>>

\vfill
\textbf{\sffamily Abstract: }\textbf{Background.} Range maps are a
useful tool to describe the spatial distribution of species. However,
they need to be used with caution, as they essentially represent a rough
approximation of a species' suitable habitats. When stacked together,
the resulting communities in each grid cell may not always be realistic,
especially when species interactions are taken into account. Here we
show the extent of the mismatch between range maps, provided by the
International Union for Conservation of Nature (IUCN), and species
interactions data. More precisely, we show that local networks built
from those stacked range maps often yield unrealistic communities, where
species of higher trophic levels are completely disconnected from
primary producers. \textbf{Methodology.} We used the well-described
Serengeti food web of mammals and plants as our case study, and provide
updated range maps for all predators by taking into account food-web
structure. We then used occurrence data from the Global Biodiversity
Information Facility (GBIF) to investigate where data is most lacking.
\textbf{Results.} We found that most predator ranges comprised large
areas without any overlapping distribution of their preys. However, many
of these areas contained GBIF occurrences of the predator.
\textbf{Conclusions.} Our results suggest that the mismatch between both
data sources could be due either to the lack of information about
ecological interactions or the geographical occurrence of preys. We
finally discuss general guidelines to help identify defective data among
distributions and interactions data, and we recommend this method as a
valuable way to assess whether the occurrence data that are being used,
even if incomplete, are ecologically accurate.
\vfill

@@

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors, geographic
inaccuracy, or sampling biases (Ladle and Hortal 2013; Hortal et al.
2015; Poisot et al. 2021). One way to identify - and potentially fix -
these errors is to combine many different pieces of information about
the occurrence of a species, so agreements and mismatches can emerge.
Here we suggest jointly analyzing species occurrence (range maps and
point occurrences) and ecological interactions to identify mismatches
between datasets.

@@

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (2021). These maps can be used in macroecological inferences in
the lack of more precise information (Fourcade 2016; Alhajeri and
Fourcade 2019), but it has been recommended that they are used with
caution since they tend to underestimate the distribution of species
that are not well-known (Herkt, Skidmore, and Fahr 2017), do not
represent spatial variation in species occurrence and abundance (Dallas,
Pironon, and Santini 2020), and can include inadequate areas within the
estimated range. Another source of species distribution information is
the Global Biodiversity Information Facility (GBIF), which is an online
repository of georeferenced observational records that come from various
sources, including community science programs, museum collections, and
long-term monitoring schemes. A great source of bias in these datasets
is the irregular sampling effort, with more occurrences originating from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods and
data aggregation (Poisot et al. 2020; Hortal et al. 2015). Nevertheless,
we have witnessed an increase in the availability of biodiversity data
in the last decades, including those collected through community science
projects (Callaghan et al. 2019; Pocock et al. 2015) and dedicated
databases, such as Mangal (Poisot et al. 2016). This provides an
opportunity to merge species distribution and ecological interaction
data to improve our predictions of where a species may be found across
large spatial scales (e.g., continental and global).

@@

We investigated the mismatch between savannah species ranges and
interactions in Africa (fig.~\ref{fig:richness}). These ecosystems host
a range of different species, including the well-characterized
predator-prey dynamics between iconic predators (e.g., lions, hyenas,
and leopards) and large herbivores (e.g., antelopes, wildebeests, and
zebras), as well as a range of herbivorous and carnivorous small
mammals. The Serengeti ecosystem has been extensively studied and its
food web is one of the most complete we have to date, including primary
producers identified to the species level. Here we focus on six groups
of herbivores and carnivores from the Serengeti Food Web Data Set
(Baskerville et al. 2011). These species exhibit direct antagonistic
(predator-prey) interactions with one another and are commonly found
across savannah ecosystems on the African continent (McNaughton 1992).
Plants in the network were included indirectly in our analyses as we do
not expect the primary producers to significantly influence the range of
herbivores for several reasons. Firstly, many savannah plants are
functionally similar (i.e., grasses, trees and shrubs) and cooccur
across the same habitats (Baskerville et al. 2011). Secondly, herbivores
in the network are broadly generalists feeding on a wide range of
different plants across habitats. Indeed, out of 129 plants in our
dataset, herbivores (n = 23) had a mean out degree (mean number of
preys) of around 22 (std = 17.5). There is also an absence of global
range maps for many plant species (Daru 2020), which prevents their
direct inclusion in our analysis. Therefore, we assume that plants
consumed by herbivores are present across their ranges, and as such the
ranges of herbivores are not expected to be significantly constrained by
the availability of food plants.

From the wider ecological network presented in Baskerville (2011), we
sampled interaction data for herbivores and carnivores. This subnetwork
contained 32 taxa (23 herbivores and 9 carnivores) and 84 interactions
and had a connectance of 0.08. Although self-loops are informative, we
removed these interactions to allow for the original IUCN ranges of
predators with cannibalistic interactions to be adjusted. We treated
this overall network as a metaweb since it \emph{should} contain all
potential species interactions between mammalian taxa occurring across
savannah ecosystems such as the Serengeti.

We compiled IUCN range maps for the 32 species included in the metaweb
from the Spatial Data Download portal
(www.iucnredlist.org/resources/spatial-data-download), which we
rasterized at 10 arc-minute resolution (\textasciitilde19 km² at the
equator). We then combined interaction data from the metaweb and
cooccurrence data generated from species ranges to create networks for
each raster pixel. This generated a total of 84,244 pixel-level
networks. These networks describe potential predation, not actual
interactions: the former is derived information from the metaweb, and
the latter is contingent on the presence of herbivores.

@@

We calculated the geographical overlap, i.e.~the extent to which
interacting predator and prey species co-occurred across their ranges,
as \(a/(a + c)\), where \(a\) is the number of pixels where predator and
prey cooccur and \(c\) is the number of pixels where only the focal
species occur. This index of geographical overlap can be calculated with
prey or predators as the focal species. Values vary between 0 and 1,
with values closer to 1 indicating that there is a large overlap in the
ranges of the two species and values closer to 0 indicating low
cooccurrence across their ranges. For each predator species, we
calculated its generality to understand whether the level of trophic
specialization (i.e., number of prey items per predator) affects the
extent to which the ranges of the species were altered. One would assume
that predators with a greater number of prey taxa (i.e., a higher
generality) are less likely to have significant changes in their range
as it is more likely that at least one prey species is present across
most of their range.

@@

For each species in the dataset we collated point observation data from
GBIF (www.gbif.org), and condensed these data into pixels representing
presence or absence of the focal taxon. These data were used to validate
the range adjustments made based on species interactions (see the
previous section). To do so, we calculated the proportion of GBIF
presence pixels occurring within both the original and adjusted species
ranges. We then compared these proportions for the predators to verify
if the range adjustments removed locations with GBIF observations, hence
likely true habitats.

@@

Predators with fewer prey lose more range with our method
(fig.~\ref{fig:degree}). For instance, both \emph{Leptailurus serval}
and \emph{Canis mesomelas} have only one prey in the Serengeti food web
(tbl.~\ref{tbl:everyone}), each of them with a very small range compared
to those of their predator. This discrepancy between range sizes
promotes significant range loss. On the other hand, predators of the
genus \emph{Panthera} are some of the most connected species, and they
also lose the least proportion of their ranges. This mismatch between
predators and preys can also be a result of taxonomic disagreement
between the geographical and ecological data. Although \emph{Canis
aureus} has the same number of prey as \emph{Caracal caracal}, none of
the prey taxa of the former occurs inside its original range
(tbl.~\ref{tbl:everyone}), which results in complete range loss.

@@

There was high variation in the overlap of predator and prey ranges
(fig.~\ref{fig:geo_diss}). The high density of points on the left-hand
side of fig.~\ref{fig:geo_diss} indicates that most preys have small
ranges in comparison to those of the set of carnivores in the networks,
resulting in either low overlap between both ranges (bottom) or high
overlap of ranges because much of that of the prey is within predators'
range (top). The top-right side of the plot encompasses situations where
the ranges of both predator and prey are similar and overlapping, while
the bottom-right part of the plot represents a situation where the range
of the predator is smaller than that of its prey and much of it occurs
within the preys' range. For example, \emph{Panthera pardus} had many
preys occurring inside its range, with highly variable levels of overlap
(tbl.~\ref{tbl:everyone}). In general, species exhibited more consistent
values of prey-predator overlap, than predator-prey overlap -- indicated
by the spread of points along the x-axis, yet more restricted variation
on the y-axis (fig.~\ref{fig:geo_diss}). There was also no overall
relationship between the two metrics, or for any predator species.

@@

degrees, total original range size (in pixels), and proportion of their

@@

(\emph{Loxodonta africana}) and how \emph{Canis aureus}'s range does not

@@

The proportion of GBIF pixels in updated ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Four species showed no difference in proportion while three species
showed only small differences (proportions of 0.01 to 0.05). On the
other hand, two species, \emph{Canis mesomelas}, and \emph{Leptailurus
serval} showed very high differences, with overlaps lower by 0.548 and
0.871 respectively. For \emph{Leptailurus serval}, none of the GBIF
observations occurred in the updated range. These two species are also
the only predators with a single prey in our metaweb. Our results
delineate how a mismatch between GBIF and IUCN databases differ greatly
with small changes in herbivore species ranges, and it is somewhat
positively related to range size for predator species. Moreover, we show
that accounting for interactions does not necessarily aggravates this
dissimilarity, but it is relevant for species with little ecological
information or specialists.

@@

The jackal is a widespread taxon in northern Africa, Europe, and
Australasia, generally well adapted to local conditions due to its
largely varied diet (Tsunoda and Saito 2020; Krofel et al. 2021).
Because of that, we expected that the \emph{Canis} species in our
dataset would be the ones losing the least amount of range, with a
higher value of the proportion of GBIF pixels within their IUCN range
maps. However, the taxonomy of this group is a matter of intense
discussion, as molecular and morphological data seem to disagree in the
clustering of species and subspecies (Krofel et al. 2021; Stoyanov
2020). This debate is indeed reflected in our analysis: the GBIF
identification of the golden jackal is incompatible with the one used by
IUCN, each of them mapping its distribution in completely different
places. This led to a complete exclusion of \emph{Canis aureus} from its
original range in our analysis, despite the fact that this species has
four documented preys in our metaweb. This example illustrates how the
taxonomic, geographical and ecological data can be used to validate one
another.

Here we show that when ecological interaction data (predator-prey
interactions within food webs) are used to refine species range maps,
there are significant reductions in the IUCN range size of predatory
organisms. Despite showing the potential importance of accounting for
species interactions when estimating the range of a species, it remains
unclear the extent to which the patterns observed represent ecological
processes or a lack of data. In the following sections, we discuss the
implications of our findings, in terms of species range maps,
interaction data, and the next steps required to enhance understanding
of species distributions using information on ecological networks.

@@

In the Serengeti food web there is a positive relationship between the
out degrees of predators and the size of their ranges
(tbl.~\ref{tbl:everyone}). In addition, our results show that there is a
negative relationship between the relative loss of predators' ranges and
their number of preys (fig.~\ref{fig:degree}), reinforcing the idea that
generalist species can preserve their distributions longer while losing
interactions. The factors limiting the geographical range of a species
in a community can vary with connectivity and richness (Svenning et al.
2014). Younger communities may be more affected by environmental
limitations because they are dominated by generalist species, while
older metacommunities are probably affected in different ways in the
center of the distribution, at the edge of ranges, and in sink and
source communities (Svenning et al. 2014; Godsoe et al. 2017; Cazelles
et al. 2016; Bullock et al. 2000). Additionally, it is likely that
species with larger ranges of distribution and those that are more
generalists would co-occur with a greater number of other species
(Dáttilo et al. 2020), while dispersal capacity of competitive species
modulate their aggregation in space and the effect of interactions on
their range limits (Godsoe et al. 2017).

@@

The geographical mismatch between predators and preys has ecological
consequences such as loss of ecosystem functioning and extinction of
populations (Anderson et al. 2016; Dáttilo and Rico-Gray 2018; Pringle
et al. 2016; Young et al. 2013). Climate change is one of the causes of
this, leading, for instance, to the decrease of plant populations due to
the lack of pollination (Bullock et al. 2000; Afkhami, McIntyre, and
Strauss 2014; Godsoe et al. 2017). However, this mismatch can also be
purely informational. When the distribution of predators and preys does
not superpose, it can mean we lack information about the distribution of
either species or about their interactions (e.g., predators may be
feeding on different species than the ones in our dataset outside the
Serengeti ecosystem). Here we addressed part of this problem by
comparing the IUCN range maps with GBIF occurrences, which helped us
clarify what is the shortfall for each species.

The lack of superposition between IUCN range maps and GBIF occurrences
suggests that we certainly do miss geographical information about the
distribution of a certain species, but it is not an indicator of the
completeness of the information about ecological interactions. However,
if both GBIF and IUCN occurrences tend to superpose and still the
species is locally removed, this indicates we don't have information
about all its interactions. The combination of this rationale with our
method of updating range maps based on ecological interactions allows us
to have a clearer idea of which information we are missing. For example,
the lion (\emph{Panthera leo}) was one of the species with the smallest
difference between the original and the updated ranges
(fig.~\ref{fig:degree}), but 59.5\% of the GBIF occurrences for this
species fell outside the IUCN range (fig.~\ref{fig:gbif}). In this
particular case, the IUCN maps seem to agree with species interaction
data. However, the disagreement between the IUCN and the GBIF databases
is concerning and suggests that the IUCN maps might underestimate the
lion's distribution. On the other hand, \emph{Leptailurus serval} and
\emph{Canis mesomelas} are two of the three species that lose the higher
proportion of range due to the lack of paths to a herbivore
(fig.~\ref{fig:degree}), but are also some of the species with the
higher proportion of GBIF occurrences inside IUCN range maps
(fig.~\ref{fig:gbif}). This indicates that the information we are
missing for these two species is related to either the occurrence of an
interaction or the presence of interacting species. To illustrate that,
we mapped the GBIF data for the prey of \emph{Leptailurus serval}, with
a mobility buffer around each point (fig.~\ref{fig:serval}). When
considering GBIF data, approximately 53\% of the prey's occurrences are
within the portion of the serval's range that was lost. With the buffer
area, this corresponds to 13\% of the lost range. This means that by
adding GBIF information, we would reduce the loss of range (or
information) for the predator by 13\% since its distribution is
conditional on the occurrence of its preys.

\begin{figure}
\hypertarget{fig:serval}{%
\centering
\includegraphics{figures/serval_mismatch_combined.png}
\caption{Mismatch between serval's range loss and GBIF occurrence of its
prey. The left panel shows the reduction of serval's range when we
consider the IUCN data on its prey. On the right panel, we added GBIF
data on both serval and its prey, with a buffer for the prey to account
for species mobility.}\label{fig:serval}
}
\end{figure}

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. GH,
FB, GD, and NF are funded by the NSERC Computational Biodiversity
Science and Services (BIOS\(^2\)) CREATE program; FB, NF, and TP are
funded by the Institute for Data Valorization (IVADO); NF and TP are
funded by a donation from the Courtois Foundation; GD is funded by the
FRQNT doctoral scholarship; TP is funded by the Canadian Institute of
Ecology \& Evolution; FW is funded by the Royal Society (Grant number:
CHL\textbackslash R1\textbackslash180156).
